Mg Depleted and K Enriched Stars from LAMOST Spectral Data

### 2018-01-01 article/ms.tex
Added intro thats way too long and focuses almost entirely on work on NGC 2419.

This is a simple template for authors to write new MNRAS papers.
See \texttt{mnras\_sample.tex} for a more complex example, and \texttt{mnras\_guide.tex}
for a full user guide.

All papers should start with an Introduction section, which sets the work
in context, cites relevant earlier studies in the field by \citet{Others2013},
and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.

>>>>

NGC 2419 is the Milky Way's third most massive cluster, and one of the oldest at an age of 12.3 +/- 1.3 Gy (CITE FORBES AND BRIDGES 2010). It resides in the outer halo 87.5 kpc (original source unknown) from the galactic centre, and has been identified as perhaps one of the most unusual clusters in the galaxy.

A spectroscopic study of RGB giant stars in NGC 2419 led to the discovery discovery of a strong anti-correlation between Mg and K in around 30-40 \% of the studied stars, along with weaker relations in Si, Sc, Ca, Ti and V (CITE MUCIARELLI AND COHEN AND KIRBY 2012). Depletions of Mg within globular clusters are not uncommon, but they are always accompanied with an increase in [Fe/H], with the mechanism for this phenomena being the introduction of $\alpha$ element poor but iron rich material released from Type Ia supernovae locally polluting the 'normal' [Mg/Fe] material released from Type II supernovae. However, the two sub-populations are indistinguishable in their [Fe/H] abundance (Cite Kirby and Cohen), and this mechanism fails to account for the enhanced K abundance or trends in the other elements.

Recently (cite Young Wook Lee 2017) used a Ca filter to identify two population in NGC 2419 corresponding to two generations of stars: G1, first generation metal poor stars, and G2, second generation stars displaying enhanced Ca and significantly increased He abundance ($\Delta Y = 0.19$). The Mg depleted population identified previously was found to be contained with the G2 population.

With the exception of NGC 2808, where several K enhanced (with weaker enhancements than NGC2419) have been found (cite Muciarelli 2015), this anti-correlation between Mg and K has, with the exception of this study, been found to be contained within NGC 2419, with a targeted search looking at the K abundance in RGB stars from clusters NGC 6752, NGC 6121, NGC 1904, and ω Cen as well as some field stars finding in all cases K abundances falling within the range of the Mg normal population in NGC 2419 (cite carreta 2013).

At present, there is no accepted nuclear process that can explain the elemental abundance in NGC 2419. Although the fact that Mg-K anti-correlation is confined to this specific globular cluster would seem to imply either a small population  of unusual polluter stars of some sort, or a single extremely massive polluter star, the details of these polluter candidates remain unknown.

An early modelling attempt at replicating the anomalous abundances in NGC 2419 assumed hot-bottom burning in AGB and super-AGB stars, with high temperatures of approximately 150 MK, was able to reproduce the Mg-K anti-correlation, through proton capture on Argon, but only if the reaction cross section was set to 100 times the standard rate or temperatures could be made higher than currently accepted values [cite VENTURA 2012].

A more recent, comprehensive attempt to replicate not only the Mg and K abundances in NGC 2419 but also the abundances of Si, Sc, Ca, Ti and V, with the aim of constraining the temperatures and pressures required was undertaken by [Cite Illiadis]. It was found that (with the exception of extremely high densities $\>10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^-4$ and 10$^8$ g/cm$^3$. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential poluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. Estimates based on current novae frequency in globular clusters by [CITE HENZE 2013] carried out in the work conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

However, the question remains: why is this signature unique to 2419? It has been suggested that the cluster's position and size, being both very massive and very distant from the Milky Way for a globular cluster, aids it in retaining material from explosive events. This would cause in its first generation polluter stars to have an increased effect on the second generation of stars. It has also been proposed that NGC 2419 is the stripped core of a dwarf spheroidal galaxy similar to $\omega$ Centauri, which would imply that in the past it was even more massive, allowing it to retain its ejecta even more efficiently.

Although to the knowledge of the author not formally studied in terms of feasibility, another suggested polluter candidate is a pair instability super nova (PISN) (CITE CARRETA 2013). Unique to extremely massive population III stars, these events involve the total destruction of the star, with no black hole remnant being left behind. The main argument for this idea is that the extreme rarity of these events, coupled with the huge masses of processed material involved, could perhaps allow the entire signature in NGC 2419 to be the result of a single event, explaining why it is not seen in any other globular clusters.

This study, which queries 450000ish LAMOST giants for depleted Mg and enhanced K, seeks to provide additional data points from which theories regarding the origin of this signature can be developed. 190 strong candidates for this signature, almost all of them not at this time associated with any globular cluster, are detailed in Table [REFERENCE TABLE HERE]. These stars occur at a wide range of metalicities, including solar metalicity. The implications for theories attempting to explain the origin of this signiature first found in NGC 2419 is likely to be significant, as for such a theory to be complete, it must now account not only for the stars in the metal poor NGC 2419, but also these stars existing at a range of metalicities in the galactic plane. Further, the task of explaining the uniqueness of NGC 2419 is not removed, as it remains the only globular cluster to significantly display this phenomena.
